---
title: "D.crab_MLE"
author: "Montana McLeod"
date: "10/20/2021"
output: html_document
---

```{r setup, include=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/Box Sync/Eder_Dungeness_2020/Model_code")
```

# load librarys
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(tidyr)
library(lubridate)
library(reshape2)
library(patchwork)
library(bbmle)
library(stats)
 library(viridis)
suppressPackageStartupMessages("Done")

```

# DeltaT function-  for modifying df to have time diff.
```{r} 
delta.T <- function(dat_total){
  
  # account for time differences
  Delta.T <- dat_total %>% 
    group_by(Date, Month, Year) %>% 
    summarise()
  
  diff.time <- NULL
  for(d in 2:nrow(Delta.T)){
    # calculate the diff. in time from the previous sample of SAME year
    if(Delta.T$Year[d] == Delta.T$Year[d-1]){
    diff.time[d] <- difftime(Delta.T$Date[d], Delta.T$Date[d-1], units = "weeks")
    }
    
    #check for if the last two samples are not the same year
    if(Delta.T$Year[nrow(Delta.T)] != Delta.T$Year[c(nrow(Delta.T) -1)]){
      diff.time[nrow(Delta.T)] = NA
    }
  }
  
  Delta.T$DeltaT <- diff.time
 
  return(Delta.T)
}
```

# Format.size.class  fxn - prepare data to be specific pots of each sample date
```{r}

format.size.class <- function(dat_total, shell){
  # NO shell-condition data avail.
  if(is.na(shell)){ 
  Pot_data <- dat_total %>% 
      group_by(Date, Month, Year, Pot) %>% 
      summarise()
  }
  # for shell-condition based analysis
  if(!is.na(shell)){ 
  Pot_data <- dat_total %>% 
      group_by(Date, Month, Year, ShellCondition, Pot) %>% 
      summarise()
  }
  
  
    # account for time between samples 
       Delta.T <- delta.T(dat_total)
       
    Pot_data <- left_join(Pot_data, Delta.T, by=c('Date','Month', 'Year'))
    ##counts the number of juveniles per pot & takes mean 
    # juvs <- Molt_szn_data %>% 
    #     group_by(Date, Month, Year, Pot) %>% 
    #   summarise(Mu_juv = count(CW < 140))
    
    juvs <- dat_total %>% 
      filter( CW < 140) %>% 
      group_by(Date, Month, Pot, Year) %>% 
      summarise(Molt.prb_juv = mean(Molt.prob),
                Mu_juv = n())
    
    #juvs <- left_join(juvs, juvs2, by=c('Date','Month', 'Year', 'Pot'))
    #juvs$Molt.prb_juv <- ifelse(is.na(juvs$Molt.prb_juv), 0, juvs$Molt.prb_juv)
               
    
    ##counts the number of sub-legals [btw 140:149] per pot & takes mean 
    subs1 <- dat_total %>% 
      filter( CW %in% 140:149) %>% 
      group_by(Date, Month, Year, Pot) %>% 
      summarise(Molt.prb_sub1= mean(Molt.prob),
                Mu_sub1 = n())
      
    
    ##counts the number of sub-legals [btw 140:149] per pot & takes mean 
    subs2 <- dat_total %>% 
      filter( CW %in% 150:159) %>% 
      group_by(Date, Month, Year, Pot) %>% 
      summarise(Molt.prb_sub2 = mean(Molt.prob),
                Mu_sub2 = n())
    
    
    ##counts the number of sub-legals per pot & takes mean 
    legals <- dat_total %>% 
      filter( CW > 159) %>% 
      group_by(Date, Month, Year, Pot) %>% 
      summarise(Molt.prb_legal = mean(Molt.prob),
                Mu_legal = n())
    
    
    #merge data
    Pot_data <- left_join(Pot_data, juvs, by=c('Date','Month', 'Year', 'Pot'))
    Pot_data <- left_join(Pot_data, subs1, by=c('Date','Month', 'Year', 'Pot'))
    Pot_data <- left_join(Pot_data, subs2, by=c('Date','Month', 'Year', 'Pot'))
    Pot_data <- left_join(Pot_data, legals, by=c('Date','Month', 'Year', 'Pot'))
    #Pot_data[is.na(Pot_data)] <- 0 
  # if we do not want the molt.prob to be zero & instead excluded, keep as na and just change sum to include na.rm = 0 for molt.in/out
 
    return(Pot_data)
    }
```

# format.dat.mll -fxn to reformat data as prior/post dataframe per year at each site 
```{r}

format.dat.mll <- function(dat_total, pre_time, post_time, pre_shell, post_shell, shell_mth){
  
   years <- unique(dat_total$Year)
   dat <- as.list(years)
   names(dat) <- c(years)
  
for(y in 1:length(years)){
  dat[[y]] <- list(prior = NULL, 
                    post = NULL)
  
  # make dfs of pre & post molt periods
 if(shell_mth == TRUE){
  prior <- dat_total[dat_total$Year == years[y] & dat_total$ShellCondition %in% pre_shell & dat_total$Month %in% pre_time, ] # pre-molt size
   prior  <- prior[!is.na(prior$Mu_sub1) | !is.na(prior$Mu_sub2), ] # must have at least one sub-leagal szied male from either sub1 or sub2 class
  post <- dat_total[dat_total$Year == years[y] & dat_total$ShellCondition %in% post_shell & dat_total$Month %in% post_time & !is.na(dat_total$Mu_legal), ] # post-molt size
 }
  
  # set months instead of frequency derived months
 if(shell_mth == FALSE){
 prior <- dat_total[dat_total$Year == years[y] & dat_total$ShellCondition %in% pre_shell & dat_total$Month %in% months[4:6], ] # pre-molt ize
  prior  <- prior  <- prior[!is.na(prior$Mu_sub1) | !is.na(prior$Mu_sub2), ] # must have at least one sub-leagal szied male from either sub1 or sub2 class
 post <- dat_total[dat_total$Year == years[y] & dat_total$ShellCondition %in% post_shell & dat_total$Month %in% months[7:9] & !is.na(dat_total$Mu_legal), ] # post-molt size
 }
  
  #exclude years that do not have enough data for the calculation 
    if(nrow(prior) > 0 & nrow(post) > 0){
      dat[[y]]$prior <- prior
      dat[[y]]$post <- post
    }
  }
return(dat)
}
```


## WDFW Data -> make into juv, mu, legal columns
```{r}
load("../Data/WDFW_total.Rdata")
sites <- unique(WDFW_pots_total$Site)

shell <- c(1,2,6:8) # shell conditions we want to reference
WDFW_total <- WDFW_pots_total[WDFW_pots_total$ShellCondition %in% shell,]

# empty list for WDFW sites
WDFW_list <- vector(mode = "list", length = length(sites))
names(WDFW_list) <- sites

WDFW_GCove <- WDFW_total[WDFW_total$Site == sites[1], ]
  WDFW_list[[1]]  <- as.data.frame(format.size.class(WDFW_GCove, T))

WDFW_KPt <- WDFW_total[WDFW_total$Site == sites[2], ]
  WDFW_list[[2]] <- format.size.class(WDFW_KPt, T)
  
WDFW_KBay <- WDFW_total[WDFW_total$Site == sites[3], ]
  WDFW_list[[3]] <- format.size.class(WDFW_KBay, T)

WDFW_Lil <- WDFW_total[WDFW_total$Site == sites[4], ]
  WDFW_list[[4]] <- format.size.class(WDFW_Lil, T)

WDFW_OBay <- WDFW_total[WDFW_total$Site == sites[5], ]
  WDFW_list[[5]] <- format.size.class(WDFW_OBay, T)

WDFW_PGam <- WDFW_total[WDFW_total$Site == sites[6], ]
  WDFW_list[[6]] <- format.size.class(WDFW_PGam, T)

WDFW_Quil <- WDFW_total[WDFW_total$Site == sites[7], ]
  WDFW_list[[7]] <- format.size.class(WDFW_Quil, T)

WDFW_SBch <- WDFW_total[WDFW_total$Site == sites[8], ]
  WDFW_list[[8]] <- format.size.class(WDFW_SBch, T)

WDFW_SHbr <- WDFW_total[WDFW_total$Site == sites[9], ]
  WDFW_list[[9]] <- format.size.class(WDFW_SHbr, T)

WDFW_Tyua <- WDFW_total[WDFW_total$Site == sites[10], ]
  WDFW_list[[10]] <- format.size.class(WDFW_Tyua, T)

WDFW_Vland <- WDFW_total[WDFW_total$Site == sites[11], ]
  WDFW_list[[11]] <- format.size.class(WDFW_Vland, T)
  
```
 
 
#freq.dat - function for automating when the pre-molt month seq is
```{r}
 months <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

freq.dat <-function(x, post_shell){
  months <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
  # create df for the combo of months that have the most legals w/ soft shell count 
  Freq_df <- data.frame(Trial = rep(NA, 12), 
                        Count = rep(NA, 12))
  Freq_df$Trial[1] <- c("January, February, March")
  Freq_df$Trial[2] <- c("February, March, April")
  Freq_df$Trial[3] <- c("March, April, May")
  Freq_df$Trial[4] <- c("April, May, June")
  Freq_df$Trial[5] <- c("May, June, July")
  Freq_df$Trial[6] <- c("June, July, August")
  Freq_df$Trial[7] <- c("July, August, September")
  Freq_df$Trial[8] <- c("August, September, October")
  Freq_df$Trial[9] <- c("September, October, November")
  Freq_df$Trial[10] <- c("October, November, December")
  Freq_df$Trial[11] <- c("November, December, January")
  Freq_df$Trial[12] <- c("December, January, February")
  
  # df of soft shell legals
  x <- x[x$CW >= 160 & x$ShellCondition %in% post_shell, ]
  
  # finding the maximum focal pt count
  for(m in 1:12){ 
    if(m == 11){
       xx  <- nrow(x[x$Month %in% months[c(m:c(m+1), 1)], ])
    }
    if(m ==12){
       xx  <- nrow(x[x$Month %in% months[c(m,1:2)], ])
    }
     xx  <- nrow(x[x$Month %in% months[m:c(m+2)], ])
  Freq_df$Count[m] <- xx
  }
return( Freq_df[which.max(Freq_df$Count), 1])
}

```



# Data reformated with the new format function by (freq. month)
```{r}
  
### data with mle format  
  WDFW_mll_list <- vector(mode = "list", length = length(sites))
names(WDFW_mll_list) <- sites

# find the most frequent consecutive 3-mo of soft shell legals - set as post values, backtrack for prior months
  freq.dat(WDFW_GCove, c(6:8))
   WDFW_mll_list[[1]] <- format.dat.mll(WDFW_list[[1]], months[c(12, 1:2)], months[3:5], c(1:2), c(6:8), T)
   freq.dat(WDFW_KBay, c(6:8))
   WDFW_mll_list[[2]] <- format.dat.mll(WDFW_list[[2]], months[c(12, 1:2)], months[3:5], c(1:2), c(6:8), T)
   freq.dat(WDFW_KPt, c(6:8))
   WDFW_mll_list[[3]] <- format.dat.mll(WDFW_list[[3]], months[1:3], months[4:6], c(1:2), c(6:8), T)
   freq.dat(WDFW_Lil, c(6:8))
   WDFW_mll_list[[4]] <- format.dat.mll(WDFW_list[[4]], months[c(11:12, 1)], months[2:4], c(1:2), c(6:8), T)
   freq.dat(WDFW_OBay, c(6:8))
   WDFW_mll_list[[5]] <- format.dat.mll(WDFW_list[[5]], months[1:3], months[4:6], c(1:2), c(6:8), T)
   freq.dat(WDFW_PGam, c(6:8))
   WDFW_mll_list[[6]] <- format.dat.mll(WDFW_list[[6]], months[1:3], months[4:6], c(1:2), c(6:8), T)
   freq.dat(WDFW_Quil, c(6:8))
   WDFW_mll_list[[7]] <- format.dat.mll(WDFW_list[[7]], months[5:7], months[8:10], c(1:2), c(6:8), T)
  freq.dat(WDFW_SBch, c(6:8))
   WDFW_mll_list[[8]] <- format.dat.mll(WDFW_list[[8]], months[5:7], months[8:10], c(1:2), c(6:8), T)
   freq.dat(WDFW_SHbr, c(6:8))
   WDFW_mll_list[[9]] <- format.dat.mll(WDFW_list[[9]], months[1:3], months[4:6], c(1:2), c(6:8), T)
   freq.dat(WDFW_Tyua, c(6:8))
   WDFW_mll_list[[10]] <- format.dat.mll(WDFW_list[[10]], months[c(12, 1:2)], months[3:5], c(1:2), c(6:8), T)
   freq.dat(WDFW_Vland, c(6:8))
   WDFW_mll_list[[11]] <- format.dat.mll(WDFW_list[[11]], months[1:3], months[4:6], c(1:2), c(6:8), T)
   
  
```


# Molt.ll.shell Fxn - for WDFW by site and shell condition 
```{r}
molt.LL.shell <- function(dat, dat_total, k1, k2, k3, Mu1, Mu2, M){
  years <- unique(dat_total$Year)
  
LL1 <- rep(NA, length(years))
LL2 <- rep(NA, length(years))
LL3 <- rep(NA, length(years))
    # prep data for the specific year 

  for(y in 1:length(years)){
     # create while loop to only work for years that has data for prior AND post
    if(!is.null(dat[[y]]$prior) & !is.null(dat[[y]]$post)){
      
    # make dfs of pre & post molt periods
    prior <- dat[[y]][1]$prior
      #dat[dat$Year == years[y] & dat$ShellCondition %in% 1:2 & dat$Month %in% c("March", "April", "May", "June"), ] # pre-molt size
    # prior <- #prior[!is.na(prior$Mu_sub1) | !is.na(prior$Mu_sub2), ] # must have at least one sub-leagal szied male from either sub1 or sub2 class
    post <- dat[[y]][2]$post
       #dat[dat$Year == years[y] & dat$ShellCondition %in% 6:8 & dat$Month %in% c("September", "October", "November", "December") & !is.na(dat$Mu_legal), ] # post-molt size
    
    # calcuate the avg time difference btw post & pre molt periods
    DeltaT =  as.numeric(abs(difftime(median(post$Date), median(prior$Date), units = "weeks" ))) / 52
    
  # BOTH sub-legal classes as Log likelihoods at time 1 [ pre-molt]
   LL1[y] <- -mean(dnbinom(x = na.omit(prior$Mu_sub1), mu = exp(Mu1), size = exp(k1), log = T))
   LL2[y] <- -mean(dnbinom(x = na.omit(prior$Mu_sub2), mu = exp(Mu2), size = exp(k2), log = T))
   
  # account for crabs molting in & out 
      lin_mu_1 = exp(Mu1) * mean(prior$Molt.prb_sub1, na.rm = T) * exp(-exp(M) * DeltaT) 
      lin_mu_2 = exp(Mu2) * mean(prior$Molt.prb_sub2, na.rm = T) * exp(-exp(M) * DeltaT)
      if(is.nan(lin_mu_1)){lin_mu_1 = 0}
      if(is.nan(lin_mu_2)){lin_mu_2 = 0}
     mu_post = lin_mu_1 + lin_mu_2
       
  # sub legal males as log likelihood at time 2 [post-molt]
    LL3[y] <- -mean(dnbinom(x = na.omit(post$Mu_legal), mu = mu_post, size = exp(k3), log = T))
    } #end of if statement - skip null years
  } #end of for loop over years
 
#if sub-legals1 or 2 have only NAs make 0
  if(is.nan(mean(LL1, na.rm = T))){LL1 = 0}
  if(is.nan(mean(LL2, na.rm = T))){LL2 = 0}

  #sum_log_lik <- sum(LL1,LL2,LL3, na.rm = T)
  sum_log_lik <- sum(mean(LL1,na.rm = T), mean(LL2, na.rm = T), mean(LL3, na.rm = T))
 sum_log_lik[which(!is.finite(sum_log_lik))] <- log(1e-320) 
 return(sum_log_lik)
}

#zz<-molt.LL.shell(wdfw_1, WDFW_list[[1]], log(10), log(10), log(10), log(mean(WDFW_list[[1]]$Mu_sub1, na.rm = T)), log(mean(WDFW_list[[1]]$Mu_sub2, na.rm = T)), log(0.2))
```





#shell condition MLE Results - WDFW data
```{r}
WDFW_Results <- data.frame(Site = sites,
                           M = rep(-999, length(sites)),
                           Std_error = rep(-999, length(sites)))

for(d in 1:length(sites)){
  
  
 Molt.LL2 <- mle2(minuslogl =  molt.LL.shell, 
                 start = list(k1 = log(10),
                                #log(mom(mean(WDFW_list[[d]]$Mu_sub1, na.rm = T), var(WDFW_list[[d]]$Mu_sub1, na.rm = T))), # lognormal params
                k2 = log(10),
                  #mom(mean(WDFW_list[[d]]$Mu_sub2, na.rm = T), var(WDFW_list[[d]]$Mu_sub2, na.rm = T))),
                k3 = log(10),
                  #mom(mean(WDFW_list[[d]]$Mu_legal, na.rm = T), var(WDFW_list[[d]]$Mu_legal, na.rm = T))),
                M = log(0.3),
                Mu1 = log(mean(WDFW_list[[d]]$Mu_sub1, na.rm = T)),
                Mu2 = log(mean(WDFW_list[[d]]$Mu_sub2, na.rm = T))), 
                data = list(dat = WDFW_mll_list[[d]], dat_total = WDFW_list[[d]]))
               #method = "L-BFGS-B", 
               #lower = rep(-1e10, 6), upper = rep(1e10, 6),
               #control = list(trace = 5, fnscale = -1) )
                #method = "Nelder-Mead") 
 print(summary(Molt.LL2)) #watch updates
 # EXP the mortality estimate 
 
 WDFW_Results[d, 2] <- coef(summary(Molt.LL2))[6,1]
 WDFW_Results[d, 3] <- coef(summary(Molt.LL2))[6,2]
}

 # calculate upper-lower cof. bounds in log space --> exp. 
 #WDFW_Results[d, 3] <- if(!is.na(coef(summary(Molt.LL2))[6,2])){
 #      coef(proffun(Molt.LL2, std.err = coef(summary(Molt.LL2))[,2], try_harder = T )[6])
 #  # want to record NA so it dpesnt no disrupt est.
 #  }else{(is.na(coef(summary(Molt.LL2))[6,2]))
 #    NA
 #  }

#}
WDFW_Results$Upper <- exp(WDFW_Results$M  + 1.96*WDFW_Results$Std_error)
  #WDFW_Results$Std_error * 1.96 + WDFW_Results$M
WDFW_Results$Lower <- exp(WDFW_Results$M  - 1.96*WDFW_Results$Std_error)
  #WDFW_Results$Std_error * 1.96 - WDFW_Results$M  
WDFW_Results$M <- exp(WDFW_Results$M)

# plot of each site with estimate of mortality 
 ggplot(WDFW_Results, aes(x=Site, y= M, color= Site)) +
  geom_errorbar(aes(ymin= Lower, ymax=Upper), width=.1) +
   ylab("Natural Mortality Rate")+
   ggtitle("WDFW Datasets")+
   ylim(0,4)+
  geom_point() +
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
   scale_color_viridis(discrete = T) +
  guides(color=FALSE)
 ### NOTES: 
 ####  sites with very low M have the most soft-shell legal sized crabs during april-June which is also the months with the highest counts for all male crabs, so the pre-molt stage is more data deficient than the post-molt stage skewing the results to a smaller mortality rate
 ### tahyua is estimated as a very large M - likely due to data having more timeseries avialable (6) than others and an over all higher avg sub-legals caught per trap --> smaller overdispersion parameter 
 
```


### bootstrap 
```{r}
set.seed(105)

store.matrixBedAngII = foreach(i = 1:2, .combine = "rbind", .packages = c("deSolve", "bbmle")) %dopar% {
  
  for(d in 1:length(sites)){
  
     index = sample(1:nrow(WDFW_list[[d]]), dim(WDFW_list[[d]])[1], replace = TRUE)
     data.all = WDFW_list[[d]][index,]
     
  
  m = mle2(minuslogl =  molt.LL.shell, 
                 start = list(k1 = log(10),
                k2 = log(10),k3 = log(10),
                M = log(0.3),
                Mu1 = log(mean(WDFW_list[[d]]$Mu_sub1, na.rm = T)),
                Mu2 = log(mean(WDFW_list[[d]]$Mu_sub2, na.rm = T))), 
                data = list(dat = WDFW_mll_list[[d]], dat_total = WDFW_list[[d]]),
           method = "Nelder-Mead",
           control = list(maxit = 1000))
  
  }
  store.matrixBedAngII[i, ] = as.vector(t(rbind(coef(summary(m))[, 1])))
}






####
d=1
wdfw_dat <- WDFW_list[[d]]
#wdfw_dat_all <- WDFW_mll_list[[d]]
years <- unique(wdfw_dat$Year)
shell = list(pre=6:8, 
            post= 1:2)

for (i in 1:1000){

  NewData <- data.frame()

  for (m in shell){
    for (y in Years){

      # resampling data 
  Howmanypots <- unique(wdfw_dat$Pot[wdfw_dat$ShellCondition %in% m  & wdfw_dat$Year== y ]) # records pot #s for each sample date
  Pots <- sample(Howmanypots, length(Howmanypots), replace=TRUE) # replacement so no duplicates


  for (p in Pots){
# create new data set and reformat for mle
    NewData.tmp <- wdfw_dat[which(wdfw_dat$Pot == p), ] 
    wdfw_mll_temp <- format.dat.mll(NewData.tmp, months[c(12, 1:2)], months[3:5], c(1:2), c(6:8), T)
####################################
    ?????? #how do we save this given that we separate it by prior/post in years?
      
    NewData = rbind(NewData, wdfw_mll_temp)
#return(NewData)
  } #end of pots loop
 } #end of years lop
} #end of shell cond loop
  
 Est_Mort_InOut <- rbind( est.mort.InOut(NewData, 2014, 1),  est.mort.InOut(NewData, 2020, 1), est.mort.InOut(NewData, 2021, 1))
Nat_Mort_InOut[,i+1] <- Est_Mort_InOut
} #end of iterations loop

```








#histograms for presentation
```{r}
 yr_2004 <- WDFW_OBay[WDFW_OBay$Year ==2004, ]
 
bl<- ggplot(yr_2004, aes(CW))+
   #geom_histogram(data= WDFW_GCove, fill = "blue", alpha = 0.2)+
    geom_histogram(data = yr_2004[yr_2004$CW %in% 140:159 & yr_2004$ShellCondition %in% 1:2,], fill = "purple", alpha = 0.4) + 
    geom_histogram(data = yr_2004[yr_2004$CW >= 160 & yr_2004$ShellCondition %in% 6:8, ], fill = "red", alpha = 0.4)
 
ble<- ggplot(yr_2004, aes(CW))+
   #geom_histogram(data= WDFW_GCove, fill = "blue", alpha = 0.2)+
    geom_histogram(data = yr_2004[yr_2004$CW%in% 140:159 & yr_2004$ShellCondition %in% 1:2& yr_2004$Month %in% months[c(12, 1:2)], ], fill = "purple", alpha = 0.4) + 
    geom_histogram(data = yr_2004[yr_2004$CW >= 160 & yr_2004$ShellCondition %in% 6:8 & yr_2004$Month %in% months[3:5], ], fill = "red", alpha = 0.6)
```
 

```{r}
WDFW_Results_setTimes <- data.frame(Site = sites,
                           M = rep(-999, length(sites)),
                           Std_error = rep(-999, length(sites)))

set_time_data <- vector("list", 11)
for(dd in 1:length(sites)){
  # data with set times of molt
  set_time_data[[dd]] <- format.dat.mll(WDFW_list[[dd]], NA, NA, F)
}


for(d in 1:length(sites)){
 Molt.LL2 <- mle2(minuslogl =  molt.LL.shell, 
                 start = list(k1 = log(10), # lognormal params
                k2 = log(10),
                k3 = log(10),
                M = log(0.3),
                Mu1 = log(mean(WDFW_list[[d]]$Mu_sub1, na.rm = T)),
                Mu2 = log(mean(WDFW_list[[d]]$Mu_sub2, na.rm = T))), 
                data = list(dat = set_time_data[[d]], dat_total = WDFW_list[[d]]),
                method="Nelder-Mead")
     
 print(summary(Molt.LL2)) #watch updates

 WDFW_Results_setTimes[d, 2:3] <- exp(coef(summary(Molt.LL2))[6,1:2])
}

# plot of each site with estimate of mortality 
ggplot(WDFW_Results_setTimes[WDFW_Results_setTimes$Site != "Squamish Hbr.", ], aes(x=Site, y= M, color= Site)) +
  geom_errorbar(aes(ymin= M-Std_error, ymax=M+Std_error), width=.1) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  guides(color=FALSE)



```



## plot of data when alternating start times for prior & post-molt cut offs
```{r}
load("../Data/WDFW_MLL_Results_StartTime.Rdata")
wdfw_mll_results$M <- exp(wdfw_mll_results$M)
wdfw_mll_results$Std_error <- exp(wdfw_mll_results$Std_error)

# plot the results by site
s1 <- ggplot(wdfw_mll_results[wdfw_mll_results$Std_error < 50, ], aes(x =start.time, y = M, color= start.time))+
  geom_errorbar(aes(ymin= M-Std_error, ymax= M+Std_error), width=.1) +
  geom_point() + 
  facet_wrap(~Site)+
  guides(color=FALSE)

s2 <- ggplot(wdfw_mll_results[wdfw_mll_results$Std_error >= 50, ], aes(x =start.time, y = M, color= start.time))+
  geom_errorbar(aes(ymin= M-Std_error, ymax= M+Std_error), width=.1) +
  geom_point() + 
  facet_wrap(~Site)+
  guides(color=FALSE)
s1 + s2

```

# count.prior.post fxn - estiamte of pre/post-molt counts by year & site
```{r}
count.prior.post <- function(site_dat){
  years <- unique(site_dat$Year)
  count_dat <- NULL
  for(y in 1:length(years)){
    x <- data.frame(year = rep(years[y], times=2),
                    Stage = c("Pre", "Post"),
                    Count = rep(NA, 2))
    # make dfs of pre & post molt periods
    x$Count[1] <- nrow(site_dat[site_dat$Year == years[y] & site_dat$CW %in% 140:159 & site_dat$ShellCondition %in% 1:2 & site_dat$Month %in% c("January","February", "March"), ]) # pre-molt size
    
    x$Count[2] <- nrow(site_dat[site_dat$Year == years[y] & site_dat$CW >=160 & site_dat$ShellCondition %in% 6:8 & site_dat$Month %in% c( "April", "May", "June"), ]) # post-molt size
    
    count_dat <- rbind(count_dat, x)
  }
  return(count_dat)
}

Gcove <- count.prior.post(WDFW_GCove)
    Gcove1 <- Gcove[Gcove$Count != 0,]
    years <- unique(Gcove1$year)
    full_yrs <- years[which(table(Gcove1$year) > 1)]
  Gcove1 <- Gcove1[Gcove1$year %in% full_yrs, ]
  Gcove1$Site = rep("GCove", nrow(Gcove1))
  
KBay <- count.prior.post(WDFW_KBay)
    KBay1 <- KBay[KBay$Count != 0,]
    years <- unique(KBay1$year)
    full_yrs <- years[which(table(KBay1$year) > 1)]
  KBay1 <- KBay1[KBay1$year %in% full_yrs, ]
  KBay1$Site = rep("KBay", nrow(KBay1))

KPt <- count.prior.post(WDFW_KPt)
    KPt1 <- KPt[KPt$Count != 0,]
    years <- unique(KPt1$year)
    full_yrs <- years[which(table(KPt1$year) > 1)]
  KPt1 <- KPt1[KPt1$year %in% full_yrs, ]
  KPt1$Site = rep("KPt", nrow(KPt1))

  Lil <- count.prior.post(WDFW_Lil)
Lil1 <- Lil[Lil$Count != 0, ]
years <- unique(Lil1$year)
full_yrs <- years[which(table(Lil1$year) > 1)]
Lil1 <- Lil1[Lil1$year %in% full_yrs, ]
Lil1$Site = rep("Lil", nrow(Lil1))

PGam <- count.prior.post(WDFW_PGam)
PGam1 <- PGam[PGam$Count != 0, ]
years <- unique(PGam1$year)
full_yrs <- years[which(table(PGam1$year) > 1)]
PGam1 <- PGam1[PGam1$year %in% full_yrs, ]
PGam1$Site = rep("PGam", nrow(PGam1))

Quil <- count.prior.post(WDFW_Quil)
Quil1 <- Quil[Quil$Count != 0, ]
years <- unique(Quil1$year)
full_yrs <- years[which(table(Quil1$year) > 1)]
Quil1 <- Quil1[Quil1$year %in% full_yrs, ]
Quil1$Site = rep("Quil", nrow(Quil1))

SHbr <- count.prior.post(WDFW_SHbr)
SHbr1 <- SHbr[SHbr$Count != 0, ]
years <- unique(SHbr1$year)
full_yrs <- years[which(table(SHbr1$year) > 1)]
SHbr1 <- SHbr1[SHbr1$year %in% full_yrs, ]
SHbr1$Site = rep("SHbr", nrow(SHbr1))

SBch <- count.prior.post(WDFW_SBch)
SBch1 <- SBch[SBch$Count != 0, ]
years <- unique(SBch1$year)
full_yrs <- years[which(table(SBch1$year) > 1)]
SBch1 <- SBch1[SBch1$year %in% full_yrs, ]
SBch1$Site = rep("SBch", nrow(SBch1))

Tyua <- count.prior.post(WDFW_Tyua)
Tyua1 <- Tyua[Tyua$Count != 0, ]
years <- unique(Tyua1$year)
full_yrs <- years[which(table(Tyua1$year) > 1)]
Tyua1 <- Tyua1[Tyua1$year %in% full_yrs, ]
Tyua1$Site = rep("Tyua", nrow(Tyua1))

Vland <- count.prior.post(WDFW_Vland)
Vland1 <- Vland[Vland$Count != 0, ]
years <- unique(Vland1$year)
full_yrs <- years[which(table(Vland1$year) > 1)]
Vland1 <- Vland1[Vland1$year %in% full_yrs, ]
Vland1$Site = rep("Vland", nrow(Vland1))

OBay <- count.prior.post(WDFW_OBay)
OBay1 <- OBay[OBay$Count != 0, ]
years <- unique(OBay1$year)
full_yrs <- years[which(table(OBay1$year) > 1)]
OBay1 <- OBay1[OBay1$year %in% full_yrs, ]
OBay1$Site = rep("OBay", nrow(OBay1))


pre_post_dat <- rbind(Gcove, KBay, KPt, SBch, Tyua, Vland, SHbr, Quil, PGam, Lil, OBay)
#pre_post_dat$Site <- rep(c("GCove", "KBay", "KPt", "SBch"), each = 40)


Pre_postNOZero_dat <- rbind(Gcove1, KBay1, KPt1, SBch1, Tyua1, Vland1, SHbr1, Quil1, PGam1, Lil1, OBay1)
```

# plots of count point data by site
```{r}
## CHANGE IT TO BE PRE-POST COLUMN AND ALL VALUES UNDER A COUNT COLUMN 
ggplot(pre_post_dat, aes(x= year, y=Count, color=Stage))+
  geom_point()+
  facet_wrap(~Site)

ggplot(Pre_postNOZero_dat[Pre_postNOZero_dat$year <= 2007, ], aes(x= year, y=Count, color=Stage))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
  facet_wrap(~Site)

```









##################### NEWPORT, OR Data
# use function to make Newport data formated by size & Sample Date
```{r}
load("../Data/Males_Molting.Rdata")

Pot_data <- format.size.class(Molt_szn_data, NA)
# catalogs dates to run through
dates <- unique(Pot_data$Date)

```


# molt.LL function - NO shell cond
```{r}

molt.LL <- function(dat, k1, k2, k3, Mu1, Mu2, M){
  months <- c("April", "August", "September")
  years <- unique(dat$Year)
  
LL1 <- rep(NA, length(years))
LL2 <- rep(NA, length(years))
LL3 <- rep(NA, length(years))
    # prep data for the specific year 

  for(y in 1:length(years)){
    prior <- dat[dat$Year == years[y] & dat$Month == months[1], ]
    post <- dat[dat$Year == years[y] & dat$Month %in% months[2:3], ]
  
  # calculate the no. of juveniles --> sub-legal for next timestep [post-molt] 
     #molt.in = sum(prior$Mu_juv, na.rm = T) * mean(prior$Molt.prb_juv, na.rm = T) 
  # calculate the no. of sublegals --> legal for next timestep [post-molt]  
    # molt.out = sum(prior$Mu_sub1, na.rm = T) * mean(prior$Molt.prb_sub1, na.rm = T) + 
             #   sum(prior$Mu_sub2, na.rm = T) * mean(prior$Molt.prb_sub2, na.rm = T)
     #molt.out = (nrow(Pre_molt_sub1) * mean(Pre_molt_sub1$Molt.prob)) + (nrow(Pre_molt_sub2) * mean(Pre_molt_sub2$Molt.prob)) 

  DeltaT =  as.numeric(abs(difftime(mean(post$Date), mean(prior$Date), units = "weeks" ))) / 52
    
  # BOTH sub-legal classes as Log likelihoods at time 1 [ pre-molt]
   LL1[y] <- -mean(dnbinom(x = na.omit(prior$Mu_sub1), mu = exp(Mu1), size = exp(k1), log = T))
   LL2[y] <- -mean(dnbinom(x = na.omit(prior$Mu_sub2), mu = exp(Mu2), size = exp(k2), log = T))
  
    # account for crabs molting in & out 
   lin_mu_1 = exp(Mu1) * mean(prior$Molt.prb_sub1, na.rm = T) * exp(-exp(M) * DeltaT) 
   lin_mu_2 = exp(Mu2) * mean(prior$Molt.prb_sub2, na.rm = T) * exp(-exp(M) * DeltaT)
      if(is.nan(lin_mu_1)){lin_mu_1 = 0}
      if(is.nan(lin_mu_2)){lin_mu_2 = 0}
     mu_post = lin_mu_1 + lin_mu_2
   
     # molt.prob is constant value for each year - avg 
      #mean(prior$Mu_sub1) * molt.in - 
     # mean(prior$Mu_sub2) * molt.out
    
  # sub legal males as log likelihood at time 2 [post-molt]
    LL3[y] <- -mean(dnbinom(x = na.omit(post$Mu_legal), mu = mu_post, size = exp(k3), log = T))
  }

#if sub-legals1 or 2 have only NAs make 0
  if(is.nan(mean(LL1, na.rm = T))){LL1 = 0}
  if(is.nan(mean(LL2, na.rm = T))){LL2 = 0}
  if(is.nan(mean(LL3, na.rm = T))){LL3 = 0}

   sum_log_lik <- sum(mean(LL1,na.rm = T), mean(LL2, na.rm = T), mean(LL3, na.rm = T))
  return(sum_log_lik )
}

#molt.LL(Pot_data, log(100),log(100), log(100), log(mean(Pot_data$Mu_sub1, na.rm = T)), log(mean(Pot_data$Mu_sub2, na.rm = T)), log(0.2))
# molt.LL(param, Pot_data, Molt_data, 2014)
# molt.LL(param, Pot_data, Molt_data, 2014)


```


# run the MLE for Newport data
```{r}
OSU_Results <- data.frame(Site = "Newport",
                           M = rep(-999, 1),
                           Std_error = rep(-999, 1))


 # neg. LL for fxn that contains pre & post-molt LL's
     Molt.LL <- mle2(minuslogl =  molt.LL, 
                 start = list(k1 = log(10), # lognormal params
                k2 = log(10),
                k3 = log(10),
                M = log(0.3),
                Mu1 = log(mean(Pot_data$Mu_sub1, na.rm = T)),
                Mu2 = log(mean(Pot_data$Mu_sub2, na.rm = T))), 
                data = list(dat = Pot_data))


 OSU_Results[1, 2] <- coef(summary(Molt.LL))[6,1]
 OSU_Results[1, 3] <- coef(summary(Molt.LL))[6,2]

OSU_Results$Upper <- exp(OSU_Results$M  + 1.96*OSU_Results$Std_error)
  #OSU_Results$Std_error * 1.96 + OSU_Results$M
OSU_Results$Lower <- exp(OSU_Results$M  - 1.96*OSU_Results$Std_error)
  #OSU_Results$Std_error * 1.96 - OSU_Results$M  
OSU_Results$M <- exp(OSU_Results$M)


# plot of each site with estimate of mortality 
 ggplot(OSU_Results, aes(x=Site, y= M, color= Site)) +
  geom_errorbar(aes(ymin= Lower, ymax=Upper), width=.1) +
   ylab("Natural Mortality Rate")+
   ylim(0,.1)+
   ggtitle("Newport Estimates")+
  geom_point() +
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
   scale_color_viridis(discrete = T) +
  guides(color=FALSE)
 
# Molt.LL1 <- mle2(minuslogl =  molt.LL, 
#                  start = list(k1 = log(10), # lognormal params
#                 k2 = log(10),
#                 k3 = log(10),
#                 M = log(0.1),
#                 Mu1 = log(mean(Pot_data$Mu_sub1, na.rm = T)),
#                 Mu2 = log(mean(Pot_data$Mu_sub2, na.rm = T))), 
#                 data = list(dat = Pot_data), 
#                 method = 'L-BFGS-B',
#            lower = rep(0, 6), upper = rep(Inf,6),
#                 control = list(trace = 5, # tracing non-negative integers
#                                fnscale= -1) ) # if fxn is negative then turns problem into amaximization problem
```





### OLD functions 

# Neg.binom fxns
```{r}

# neg-binomial fxn using mu of each pot/ sample
mtmp <- function(mu, k) {
   -sum(dnbinom(x1, mu = mu, size = k, log = TRUE))
}

# neg-binomial for predicted next timestep
mtmp_full <- function(mu, k, M) {
  mu_final <- ( mu * exp(-M) + (mu * molt.in) - (mu * molt.out) )
   -sum(dnbinom(x2, mu = mu_final, size = k, log = TRUE))
  ## CANNOT HAVE NEGATIVE MUs GO THROUGH DBINOM
  
}
```

# Run the neg.binom fxns with our data & in MLE2
```{r}
load("~/Box Sync/Eder_Dungeness_2020/Data/D.crab_MLE.Rdata")
years <- unique(Pot_data$Year)
months <- as.character(unique(Pot_data$Month))
MLE_Results_N0.prior <- NULL
MLE_Results_N0.post  <- NULL
MLE_Results_N1 <- NULL

# run the mLE2 function over different scenarios of years
  #' N0_prior : pop at pre-molt stage
  #' N0_post : pop. at post-molt stage
  #' N1: estiamted pop at next timestep w/ n0_prior mu
 for(y in 1:length(years)){
  sub_data_prior <- Pot_data[Pot_data$Year == years[y] & Pot_data$Month == months[1], ]
  sub_data_prior <- sub_data_prior[sub_data_prior$Mu_sub != 0, ] # made != 0 b/c needs a finite value 
  
  sub_data_post <- Pot_data[Pot_data$Year == years[y] & Pot_data$Month %in% months[2:3], ]
  sub_data_post <- sub_data_post[sub_data_post$Mu_sub != 0, ]

# neg. LL for N0 PRE-MOLT
  x1 <- rnbinom(n = nrow(sub_data_prior), mu = 2, size = 10)
  N0_prior <- mle2( minuslogl = mtmp, 
                 start = list(k = 1), 
                 data =  list(mu = sub_data_prior$Mu_sub) )  
  
   MLE_Results_N0.prior <- list(index = years[y], summary(N0_prior) )


 # neg. LL for N0 POST-MOLT
  x2 <- rnbinom(n = nrow(sub_data_post), mu = 2, size = 10)
  N0_post <- mle2(minuslogl = mtmp, 
                 start = list(k = 1), 
                 data = list(mu = sub_data_post$Mu_sub)) 
  MLE_Results_N0.post <- list(index = years[y], summary(N0_post) )

 
# values to calculate movement w/in size class [prior b4 post]
   molt.in = sub_data_prior$Mu_juv * sub_data_prior$Molt.prb_juv
   #molt.in <- molt.in[molt.in != 0]
   molt.out = sub_data_prior$Mu_sub * sub_data_prior$Molt.prb_sub
 
 # neg. LL for next timestep N1 - mu from m0_prior data
  N1 <- mle2(minuslogl =  mtmp_full, 
                 start = list(k = 2, M = 0.5), 
                data = list(mu = sub_data_prior$Mu_sub) )
   MLE_Results_N1 <- list(index = years[y], summary(N1) )
}

```


# plot of data
```{r}
suppressWarnings( p0 <- profile(m0) )

par(mfrow=c(1,2))
plot(p0,plot.confstr=TRUE)
```

