---
title: "Distributions and Tables"
author: "Montana McLeod"
date: "10/15/2020"
output: html_document
---

```{r setup, include=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/Documents/THESIS/Dungeness_crab/DungenessCrabMortality/Code")
```


```{r warnings=FALSE}
require(ggplot2)
require(Sleuth3)
require(multcomp)
require('patchwork')
require('RColorBrewer')
require(viridis)
require(dplyr)
require(lubridate)
require(wesanderson)
suppressPackageStartupMessages("Done")
```


# Load Manipulated Data
## Mark: can I better source multiple functions? I tried source(c()) around all R scripts
```{r}
load("../Data/Source.D.crab.Rdata")
data_Molt <- subset(Pots_N,  Month == "April" | Month == "September") #Pre-molt and post-molt
#ONLY sublegal males (9% of total crabs collected via pots)
Sub_males <- subset(Pots_N, Sex== 1 & CW >= 140 & CW <= 159, select = Date : Soak.Time.hr)

source("./Library/Total.Pots.Crabs.R")
source("./Library/Total.Crabs.Per.Pot.R")
source("./Library/CPUE.R")
source("./Library/Est.Mort.CPUE.R")
source("./Library/molt.prob.R")
```


# Boxplots of Trawl data & Pot data
```{r}
#shows TRAWL data w/ potential outliers by month
trawl <- ggplot(Trawl, aes(Month, CW))+ geom_boxplot(fill = "green")+ ggtitle("Trawl Data")
 
#shows potential outliers by month
pot <- ggplot(Pots_N, aes(Month, CW))+ geom_boxplot(fill = "blue") + ggtitle("Pot Data")
trawl + pot
```

# DISTRIBUTIONS
## Size-frequency Distribution data
### Distribution of Newport samples across Months [remaining analysis is from Newport]
```{r echo=FALSE}
#Distribution by month - yearly sampled data
  ggplot(Pots_N, aes(x=CW, fill=as.character(Year))) +
    geom_histogram(binwidth= 10)+ 
    facet_wrap(~Month, nrow=2)+ #separates the graphs by month
    scale_fill_brewer(palette = "Accent") #gradient of grays
    #guides(fill=FALSE) #removes the auto. legend 
```

# Size-frequency Distribution data
## Distribution of males caught in pots
```{r}
ggplot(Pots_N %>% group_by(Sex ==1), aes(x=CW, fill=as.character (Year))) +
    geom_histogram(binwidth= 10)+ 
    facet_wrap(~Month, nrow=2)+ #separates the graphs by month
    scale_fill_brewer(palette = "Accent") #colors of the distributions 
```

# Size-frequency Distribution data
## Distribution of sub-adult males caught in pots
```{r}
ggplot(Sub_males, aes(x=CW, fill=as.character (Year))) +
     geom_histogram(binwidth=4)+ 
     facet_wrap(~Month, nrow=2) + #separates the graphs by month
     scale_fill_brewer(palette = "Accent") +
     ggtitle("Sub-Adult males")
```


# LINEAR REGRESSION
##  Probabilty of molt using Zhang et al 2004 values
```{r}
Pots_N <-molt.prob(Pots_N)
ggplot(Pots_N, aes(x=CW, y= Molt.prob)) +
  geom_point(aes(color = CW)) +
  scale_color_viridis_c()+
  labs(title= "Size-specific molting probabilities")
```


# Shell conditions
```{r}
#Pots_N <- Pots_N[Pots_N$Date == "10/17/20" | Pots_N$Date =="9/1/20" | Pots_N$Date =="9/15/20" |Pots_N$Date == "9/29/20",]
Pots_N <- Pots_N[Pots_N$Date == "2020-10-17" | Pots_N$Date =="2020-09-01" | Pots_N$Date =="2020-09-15" |Pots_N$Date == "2020-09-29",]
X <- Total_data_N[ Total_data_N$Date == "2020-10-17" | Total_data_N$Date =="2020-09-01" | Total_data_N$Date =="2020-09-15" |Total_data_N$Date == "2020-09-29",]

# may need to change X below to shell cond 
ggplot(X, aes(x=Shell.Cond, fill= Shell.Cond))+
  geom_bar()+
  theme(axis.text.x=element_text(angle=40, hjust=1))+
  facet_wrap(~Date, nrow=2)
```



# ALL Data for sub-males 
```{r}
# just sub adults
Trawl_Sub <-size.class(Trawl, 1, "sub")
Pots_N_Sub <- size.class(Pots_N, 1, "sub")

#all datasources together 
All <- rbind(MCR_Pots_Sub[, c(2,3,5)], MCR_Seine_Sub[, c(2,3,5)], Pots_N_Sub[, c(3,2,4)],  Sub_males_Estr[, c(4,2,7)], Trawl_Sub[, c(3,2,4)] )
All$Data <- c(rep("MCR_Pots", nrow(MCR_Pots_Sub)), rep("MCR_Seine", nrow(MCR_Seine_Sub)), rep("Newp_pots", nrow(Pots_N_Sub)), rep("Yaq_trawl", nrow(Sub_males_Estr)), rep("Newp_Trawl", nrow(Trawl_Sub)) )

ggplot(All, aes(x=CW, fill=as.character (Data))) +
     geom_histogram(binwidth=4)+ 
     facet_wrap(~Month, nrow=2) + #separates the graphs by month
     scale_fill_brewer(palette = "Accent")

```



```{r}

# make them just 
MCR_Pots_Sub <- MCR_Pots_Sub[MCR_Pots_Sub$Year != 2009, ]
#No MCR seine b/c none fit the criteria
Pots_N_Sub <- Pots_N_Sub[Pots_N_Sub$Year == 2014 | Pots_N_Sub$Year == 2020, ]
Sub_males_Estr <- Sub_males_Estr[Sub_males_Estr$Year != 2004, ]
Trawl_Sub <- Trawl_Sub[Trawl_Sub$Year == 2011 | Trawl_Sub$Year == 2014, ]

#remove useless months
All_Criteria_sub <- rbind(MCR_Pots_Sub[, c(2,3,5)],  Pots_N_Sub[, c(3,2,4)],  Sub_males_Estr[, c(4,2,7)], Trawl_Sub[, c(3,2,4)] )
All_Criteria_sub$Data <- c(rep("MCR_Pots", nrow(MCR_Pots_Sub)), rep("Newp_pots", nrow(Pots_N_Sub)), rep("Yaq_trawl", nrow(Sub_males_Estr)), rep("Newp_Trawl", nrow(Trawl_Sub)) )


ggplot(All_Criteria_sub, aes(x=CW, fill=as.character (Data))) +
     geom_histogram(binwidth=4)+ 
     facet_wrap(~Month, nrow=2) + #separates the graphs by month
     scale_fill_brewer(palette = "Accent")


Yr_03 <- ggplot(All_Criteria_sub[All_Criteria_sub$Year == 2003,], aes(x=CW, fill=Data)) +
     geom_histogram(binwidth=4)+ 
     facet_wrap(~Month) + #separates the graphs by monthscale_fill_brewer("steelblue")
     scale_fill_manual(values = c("Yaq_trawl" = "steelblue"))+
      ggtitle("Year 2003")

Yr_05 <- ggplot(All_Criteria_sub[All_Criteria_sub$Year == 2005,], aes(x=CW, fill=Data)) +
     geom_histogram(binwidth=4)+ 
     facet_wrap(~Month) + #separates the graphs by month scale_fill_brewer("steelblue")
    scale_fill_manual(values = c("Yaq_trawl" = "steelblue")) +
    ggtitle("Year 2005")

Yr_08 <- ggplot(All_Criteria_sub[All_Criteria_sub$Year == 2008,], aes(x=CW, fill=Data)) +
     geom_histogram(binwidth=4)+ 
     facet_wrap(~Month) + #separates the graphs by month scale_fill_brewer("orange")
    scale_fill_manual(values = c( "MCR_Pots" = "salmon1")) +
    ggtitle("Year 2008")

Yr_11 <- ggplot(All_Criteria_sub[All_Criteria_sub$Year == 2011,], aes(x=CW, fill=Data)) +
     geom_histogram(binwidth=4)+ 
     facet_wrap(~Month) + #separates the graphs by month scale_fill_brewer("darkolivegreen")
    scale_fill_manual(values = c( "Newp_Trawl" = "yellow3")) +
    ggtitle("Year 2011")

Yr_14 <- ggplot(All_Criteria_sub[All_Criteria_sub$Year == 2014,], aes(x=CW, fill=Data)) +
     geom_histogram(binwidth=4)+ 
     facet_wrap(~Month) + #separates the graphs by month scale_fill_brewer("purple","darkolivegreen")
    scale_fill_manual(values = c("Newp_pots" = "darkgreen", "Newp_Trawl" = "yellow3"))+
    ggtitle("Year 2014")

Yr_20 <- ggplot(All_Criteria_sub[All_Criteria_sub$Year == 2020,], aes(x=CW, fill=Data)) +
     geom_histogram(binwidth=4)+ 
     facet_wrap(~Month) + #separates the graphs by month scale_fill_brewer("purple")
    scale_fill_manual(values = c("Newp_pots" = "darkgreen")) +
    ggtitle("Year 2020")

Yr_03 + Yr_05 + Yr_08 
Yr_11 + Yr_14 + Yr_20 

```

