---
title: "Distributions and Tables"
author: "Montana McLeod"
date: "10/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Documents/THESIS/Dungeness_crab/DungenessCrabMortality/Code")
```


```{r warnings=FALSE}
library(ggplot2)
library(Sleuth3)
library(multcomp)
library('patchwork')
library('RColorBrewer')
library(viridis)
library(dplyr)
library(lubridate)
library(wesanderson)
suppressPackageStartupMessages("Done")
```


# Load Manipulated Data
## Mark: can I better source multiple functions? I tried source(c()) around all R scripts
```{r}
load("../Data/Source.D.crab.Rdata")
data_Molt <- subset(Pots_N,  Month == "April" | Month == "September") #Pre-molt and post-molt
#ONLY sublegal males (9% of total crabs collected via pots)
Sub_males <- subset(Pots_N, Sex== 1 & CW >= 140 & CW <= 159, select = Date : Soak.Time.hr)

source("./Library/Total.Pots.Crabs.R")
source("./Library/Total.Crabs.Per.Pot.R")
source("./Library/CPUE.R")
source("./Library/Est.Mort.CPUE.R")
```


# Boxplots of Trawl data & Pot data
```{r}
#shows TRAWL data w/ potential outliers by month
trawl <- ggplot(Trawl, aes(Month, CW))+ geom_boxplot(fill = "green")
 
#shows potential outliers by month
pot <- ggplot(Pots_N, aes(Month, CW))+ geom_boxplot(fill = "blue")
 
trawl + pot
```

# DISTRIBUTIONS
## Size-frequency Distribution data
### Distribution of Newport samples across Months [remaining analysis is from Newport]
```{r echo=FALSE}
#Distribution by month - yearly sampled data
  ggplot(Pots_N, aes(x=CW, fill=as.character(Year))) +
    geom_histogram(binwidth= 10)+ 
    facet_wrap(~Month, nrow=2)+ #separates the graphs by month
    scale_fill_brewer(palette = "Accent") #gradient of grays
    #guides(fill=FALSE) #removes the auto. legend 
```

# Size-frequency Distribution data
## Distribution of males caught in pots
```{r}
ggplot(Pots_N %>% group_by(Sex ==1), aes(x=CW, fill=as.character (Year))) +
    geom_histogram(binwidth= 10)+ 
    facet_wrap(~Month, nrow=2)+ #separates the graphs by month
    scale_fill_brewer(palette = "Accent") #colors of the distributions 
```

# Size-frequency Distribution data
## Distribution of sub-adult males caught in pots
```{r}
ggplot(Sub_males, aes(x=CW, fill=as.character (Year))) +
     geom_histogram(binwidth=4)+ 
     facet_wrap(~Month, nrow=2) + #separates the graphs by month
     scale_fill_brewer(palette = "Accent") 
```


# LINEAR REGRESSION
##  Probabilty of molt using Zhang et al 2004 values
```{r}
Pots_N <-molt.prob(Pots_N)
ggplot(Pots_N, aes(x=CW, y= Molt.prob)) +
  geom_point(aes(color = CW)) +
  scale_color_viridis(option = "D")+
  labs(title= "Size-specific molting probabilities")
```