---
title: "Old_D.crab_code"
author: "Montana McLeod"
date: "8/31/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("D.crab.Rdata")
```

# old function for total crabs per pot
```{r}
total.pots.crabs <- function(data, size){
 #When there is a different sample date (i.e. sample), records total pots used and number of crabs collected 
 data$Total.pots <- rep(NA,nrow(data))
   data$Total.crabs <- rep(NA,nrow(data))
   
   j=1
   for(i in 2:nrow(data)){
     if(data$Date[i] !=  data$Date[i-1])
     { data$Total.pots[j:i] <- data$Pot[i-1];
     data$Total.crabs[j:i] <- (i-j);
     j=i}
  #   
     if(i == nrow(data)) 
     {data$Total.pots[j:i] <- data$Pot[i];
     data$Total.crabs[j:i] <- i-j +1}
   }
  
```


# Resampling of Our calculation [INOut]
```{r}
# Function for molt surv. 
surv.molt <- function(y){
  
#create pre and post df for each size class 
Pre_molt_juv <- subset(Resample_InOut, Carapace.Width <= 139 & Month == "April" & Year == y) #juveniles moving into sub-adults
    if(nrow(Pre_molt_juv) == 0){
      Pre_molt_juv[1,5] <- 0
    }

# sub-adults have to be spilt into two b/c of calc
Pre_molt_sub1 <- subset(Resample_InOut, Carapace.Width %in% c(140:149) & Month == "April"& Year == y)
    if(nrow(Pre_molt_sub1) == 0){
      Pre_molt_sub1[1,5] <- 0
    }

Pre_molt_sub2 <- subset(Resample_InOut, Carapace.Width %in% c(150:159) & Month == "April"& Year == y)
    if(nrow(Pre_molt_sub2) == 0){
      Pre_molt_sub2[1,5] <- 0
    }
Post_molt_sub <- subset(Resample_InOut, Carapace.Width %in% c(140:159) & Month == "September" & Year == y)
    if(nrow(Post_molt_sub) == 0){
      Post_molt_sub[1,5] <- 0
    }

#create totals for pre & post-molt to get prop. 
Pre_total <-subset(Resample_InOut, Month == "April" & Year == y)
Post_total <- subset(Resample_InOut,  Month == "September" & Year == y)


molt.in = nrow(Pre_molt_juv) * mean(Pre_molt_juv$Molt.prob) 
molt.out = (nrow(Pre_molt_sub1)* mean(Pre_molt_sub1$Molt.prob)) + (nrow(Pre_molt_sub2)* mean(Pre_molt_sub2$Molt.prob)) 
 survived <- (nrow(Post_molt_sub) + molt.in - molt.out) / ((nrow(Pre_molt_sub1) + nrow(Pre_molt_sub2)))
 return(survived)
}


# creates a new df to store the randomized data
Resample_InOut <- Ap_Sept_Data[, 1:4]
#Reorder by date
  index <- with(Resample_InOut, order(Date)) #organizes by month then year
  Resample_InOut <- Resample_InOut[index,]
  Resample_InOut$Carapace.Width <- rep(NA, nrow(Resample_InOut))

# running iterations to recieve an avg  
Nat_Mort_InOut <- data.frame(Sample_Date = c("2014-09-17","2020-09-15"))
dates <-as.data.frame(table(Ap_Sept_Data$Date)) #store total # of crabs per sample date to use in loop
dates <- dates[,2]

for(j in 1: tstep){
  #Run randomizer
    #Resample_InOut$Carapace.Width <- sample(Ap_Sept_Data$Carapace.Width, nrow(Ap_Sept_Data), replace= TRUE)
    i=1 # starting point in column
    for(d in dates){
         Resample_InOut$Carapace.Width[i:(i+d-1)]  <- sample(Ap_Sept_Data$Carapace.Width[i:(i+d-1)], d , replace= T)
         i= (i+d) #creates new starting pt
         
         #Need to limit the loop
          if(i == 348){
             break
           }
    }
    
    Resample_InOut$ Molt.prob <- rep(NA, nrow(Resample_InOut))
  #estimating molt prob. based off of zhang's equation(figure 3)
  for(i in 1:nrow(Resample_InOut)){
    if(Resample_InOut$Carapace.Width[i] <= 150)
    { Resample_InOut$Molt.prob[i] <- (-0.0014* Resample_InOut$Carapace.Width[i]) + 1.14 }
  
    if(Resample_InOut$Carapace.Width[i] > 150)
    { Resample_InOut$Molt.prob[i] <- (-0.014* Resample_InOut$Carapace.Width[i]) + 2.71 }
    
    if(Resample_InOut$Carapace.Width[i] > 193)
    { Resample_InOut$Molt.prob[i] <- 0 }
  }

  Est_Mort_InOut <- rbind(surv.molt(2014), surv.molt(2020)) 
  Nat_Mort_InOut[,j+1] <- Est_Mort_InOut
}

# Calculate variance across Resampling iterations 
Nat_Mort_InOut$Mean <- rep(NA, nrow(Nat_Mort_InOut))
Nat_Mort_InOut$Variance <- rep(NA, nrow(Nat_Mort_InOut))
for( j in 1:nrow(Nat_Mort_InOut)){
  z <- unlist(Nat_Mort_InOut[j,2:(tstep+1)])
  Nat_Mort_InOut$Mean[j] <- mean(na.omit(z))
  Nat_Mort_InOut$Variance[j] <- var(na.omit(z))
}
Nat_Mort_InOut <- melt(Nat_Mort_InOut)
###Nat_Mort_InOut <- subset(Nat_Mort_InOut, Sample_Date == "2014-09-17" |  Sample_Date == "2020-09-15")

#Nat_Mort$Real_N.Mort <-Orig_N_mort # adds column of real values based on our data
Avgs_InOut <- Nat_Mort_InOut[2001:2004,] 


ggplot(Nat_Mort_InOut[c(1:2000),], aes(Sample_Date, value))+
  geom_boxplot()+
  labs(title= "Variance of Natural Mortality Rates in Sub-legal Males", x= "Date", y= "Daily Natural Mortality Rate")

```



# General bootstrapping
```{r}
## calculate beta distrivution parameters using MOM from week 2
beta.mom<-function(mean,sd){
  v<-sd**2
  x<-mean
  a<-x*(x*(1-x)/v-1)
  b<-(1-x)*(x*(1-x)/v-1)
  c(a,b)
}
## calculate beta parameters
beta.prm<-beta.mom(molt.2014, 0.05)
## generate harvest rates from beta distrubution
sim.InOut<- rbeta(1000,beta.prm[1],beta.prm[2])

mean(rand.2014)
sd(rand.2014)
quantile(rand.2014,c(.025,0.975))


#Resampling for our calc.


```




# 2015 jan/ may CPUE check
```{r}
Jan_May_sub <- subset(Up_data_N, Year ==2015 & Sex ==1 & Carapace.Width %in% c(140:149) & Month == "January" | Month == "May", select = c(Date : Carapace.Width, Pot, Soak.Time.hr, Total.pots, Effort))

Jan_May_CPUE<- Jan_May_sub %>% 
  select(Date, Month, Year, Pot, Total.pots, Effort) %>% 
  group_by(Date, Month, Year, Pot, Total.pots, Effort) %>% 
  count(Pot)
colnames(Jan_May_CPUE)[colnames( Jan_May_CPUE) == "n"]  <- "Crabs.per.Pot"
Jan_May_CPUE <- as.data.frame(Jan_May_CPUE)

#Reorder by date
index <- with(Jan_May_CPUE, order(Date, Pot)) #organizes by month then year
Jan_May_CPUE <- Jan_May_CPUE[index,]

# run resampling throug 1000 iterations
Nat_Mort <- data.frame(Sample_Date = as.character(unique(Jan_May_CPUE$Date)))
Sim_CPUE <- Jan_May_CPUE # rename so dont overright original resample_cpue

tstep = 1000
for (j in 1:tstep){
  #Randomization of crabs per pot
  Sim_CPUE$Crabs.per.Pot  <- sample(Jan_May_CPUE$Crabs.per.Pot, nrow(Jan_May_CPUE), replace= TRUE)
  
  x <- aggregate(cbind(Total.pots, Effort) ~ Date + Month + Year, data = Sim_CPUE, max) #takes consistent values
  y <- aggregate(Crabs.per.Pot ~ Date + Month + Year, data = Sim_CPUE, sum) #takes sum of crabs
  Est_Mort_CPUE <- merge(x,y, all=TRUE) #averages by year
  
  # CPUE = # of crabs caught in a string of traps / standard effort
  Est_Mort_CPUE$CPUE <- Est_Mort_CPUE$Crabs.per.Pot / Est_Mort_CPUE$Effort
  
  # Making a Julian date by the year 
  Est_Mort_CPUE$J.date <- julian.Date(Est_Mort_CPUE$Date, as.Date("2014-01-01"), by= 'Year' )
  for (i in 1:nrow(Est_Mort_CPUE)){
    if(Est_Mort_CPUE$Year[i] != 2014) {
      n = Est_Mort_CPUE$Year[i] - 2014;
      Est_Mort_CPUE$J.date[i] <- Est_Mort_CPUE$J.date[i] - (365 * n) 
    }
  }
  Est_Mort_CPUE$N.Mort <- rep(NA, nrow(Est_Mort_CPUE)) #NULL column
  
  
  #estimating mortality for samples in same year but different months [zhang eqn]
  for(i in 2:nrow(Est_Mort_CPUE)){
    if(Est_Mort_CPUE$Year[i] == Est_Mort_CPUE$Year[i-1] & Est_Mort_CPUE$Month[i] !=   Est_Mort_CPUE$Month[i-1]) {
      Est_Mort_CPUE$N.Mort[i] <- (-1/( (Est_Mort_CPUE$J.date[i]) - (Est_Mort_CPUE$J.date[i-1]) ) )* log(Est_Mort_CPUE$CPUE[i] / Est_Mort_CPUE$CPUE[i-1] )  }
  }
  
  # store the natural mortalities for each iteration
  Nat_Mort[,j+1]  <- Est_Mort_CPUE$N.Mort
  
}

Nat_Mort <- Nat_Mort[2, ] #keeps only the estimates
#Nat_Mort[,j] <- Nat_Mort[which(is.finite(Nat_Mort$value)),j] #removes any infinite values
Nat_Mort <- do.call(data.frame,lapply(Nat_Mort, function(x) replace(x, is.infinite(x),NA)))

# Calculate variance across Resampling iterations 
Nat_Mort$Mean <- rep(NA, nrow(Nat_Mort))
Nat_Mort$Variance <- rep(NA, nrow(Nat_Mort))
for( j in 1:nrow(Nat_Mort)){
  z <- unlist(Nat_Mort[j,2:(tstep+1)])
  Nat_Mort$Mean[j] <- mean(na.omit(z))
  Nat_Mort$Variance[j] <- var(na.omit(z))
}
Nat_Mort <- melt(Nat_Mort)

#Nat_Mort$Real_N.Mort <-Orig_N_mort # adds column of real values based on our data
CPUE_avgs <- Nat_Mort[1001:1002,] 
CPUE_avgs[3:4,1] <- CPUE_avgs$Sample_Date[1:2]
CPUE_avgs[,4] <- c( rep("daily", 2), rep("annual", 2))
CPUE_avgs[3:4,2] <- CPUE_avgs$variable[1:2]
CPUE_avgs[3:4,3] <- CPUE_avgs$value[1:2] * 365
```


# resampling the brutt force way
```{r}

  empty <- c(2, 1:2, 4:6, 9:10, 4, 11) # Pots that are missing from the df - a sub-adult was not observed/ collected in every pot that was deployed
 ## Would like to do this a cleaner way 
 
# Add rows for unaccounted sub-legal males from each pot of total pots
     df <- data.frame(Date=rep(NA,length(empty)),
                      Month= c(rep(4, 8), rep(9,2) ),
                      Year= c(2014, rep(2020,(length(empty)-1))),
                      Pot= empty, 
                      Total.pots= rep(0,length(empty)), 
                      Total.crabs= rep(0,length(empty)),
                      Effort = rep(0,length(empty)),
                      Crabs.per.Pot = rep(0,length(empty)))
Resample_CPUE <- rbind(Resample_CPUE, df)

#Me poorly reorganizing to account for 0 sub-legal crabs in the pots & placing them in order
Resample_CPUE <- Resample_CPUE[c(1, 20, 2:5, 21:22, 6, 23:25, 7:8, 26:27, 9:13, 28, 14:19, 29), ]
Resample_CPUE$Date[2] <- Resample_CPUE$Date[1]
Resample_CPUE$Date[c(7:8,10:12,15:16)] <- Resample_CPUE$Date[9]
Resample_CPUE$Date[22] <- Resample_CPUE$Date[21]
Resample_CPUE$Date[29] <- Resample_CPUE$Date[28]

```



# Size-frequency Distribution data
## Modifications for the final calculations [Zhang et al 2004]
```{r}
#ONLY sublegal males (9% of total crabs collected via pots)
Sub_males <- subset(Pots_N, Sex== 1 & CW >= 140 & CW <= 159, select = Date : Total.pots)

#assigns the total number of pots per string via sample date (NT) & total crabs per string
Sub_males$Total.crabs <- rep(NA,nrow(Sub_males))
#Sub_males$Total.pots[length(Sub_males$Total.pots)] <- Sub_males$Pot[length(Sub_males$Total.pots)] 

#When there is a different sample date (i.e. sample), records total pots used and number of crabs collected 
j=1
for(i in 2:nrow(Sub_males)){
  if(Sub_males$Date[i] !=  Sub_males$Date[i-1])
  {Sub_males$Total.crabs[j:i] <- (i-j);
  j=i}
  
  if(i == nrow(Sub_males)) 
  {Sub_males$Total.crabs[j:i] <- i-j +1}
}
```


# Calculating CPUE & Estimating Mortality for Sub-legal Males [Zhang et al 2004]
```{r}
#NULL ### CPT = 14.62 * (1 - exp(-0.19 * soak_time )) 
#Sub_males$CPT <- 14.62 * (1 - exp(-0.19 * Sub_males$Soak.Time.hr ))

#NULL ### STANDARD SOAK TIME = CPT(soak time by hour) / CPT(1 hour)
#Sub_males$Sta_ST <- (1 - exp(-0.19 * Sub_males$Soak.Time.hr )) / (1 - exp(-0.19))

#NULL ### STANDARD EFFORT = standard soak time * # of traps in a sample string
#Sub_males$Sta_Effort <- Sub_males$Sta_ST * Sub_males$Total.pots

# EFFORT: = soak time * # of traps in a sample string
Sub_males$Effort <- Sub_males$Soak.Time.hr * Sub_males$Total.pots

# CPUE = # of crabs caught in a string of traps / standard effort
Sub_males$CPUE <- Sub_males$Total.crabs / Sub_males$Effort
```


# Probability of Molt  (Sub-Legal Males)
```{r}
# Before Molt
t_1 = (31 + 28 + 31 +30 + 31 +1) # AKA June 1st in julian dates
# After Molt
t1 = (t_1 + 30 + 31 + 31 + 30 ) # AKA October 1st in julian 
# Instantaneous Molt time 
t0 = t_1 + ( (t1 - t_1)/2) #AKA August 1st

# Equation for CPUE at t0
CPUE.t0 <- rep(NA, nrow(avgs))
for(i in 2:nrow(Est_Mort_sub)){
  if(Est_Mort_sub$Year[i] == Est_Mort_sub$Year[i-1] & Est_Mort_sub$Month[i-1] >= 5 & Est_Mort_sub$Month[i] != Est_Mort_sub$Month[i-1] ) #added >= 5 because need it to be close to T-1 (June) so >= May
    {CPUE.t0[i] <- Est_Mort_sub$CPUE[i-1] * exp( -1*Mean.Mort_sub * (t0 - t_1)) }
}
CPUE.t0 <- na.omit(CPUE.t0 )
P.molt <- cbind(avgs, CPUE.t0)  #adds the CPUE.t0    


#Equation for CPUE.old [old shell ind.]
with (Est_Mort_sub, table(Year, Month)) #when last sample collected by year
CPUE.old <- rep(NA, nrow(Est_Mort_sub))
for(i in 2:nrow(Est_Mort_sub)){
  if(Est_Mort_sub$Year[i] == Est_Mort_sub$Year[i-1] & Est_Mort_sub$Month[i] >= 9 & Est_Mort_sub$Month[i] != Est_Mort_sub$Month[i-1] ) #added >= 9 because need it to be close to T1 (October) so >= September
    {CPUE.old[i] <- Est_Mort_sub$CPUE[i] * exp( -1*Mean.Mort_sub * (t0 - t1)) }
}
CPUE.old <- c(na.omit(CPUE.old ),0) #NOT enough old shell sample days 
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~REMOVE THE SECOND 0^^ WHEN 2020 GETS UPDATD~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

P.molt <- P.molt[c(1:2,4), ]#removes yeare 2015 because does NOT have sufficient end month (T1)
P.molt <- cbind(P.molt , CPUE.old)  #adds the CPUE.t0    
Prob.Molt_sub <- (P.molt$CPUE.t0 - P.molt$CPUE.old) / P.molt$CPUE.t0
Prob.Molt_sub
```


## molt increments [Jamieson 1989]
### range 80 ≤ MW ≤ 174 mm
```{r}
Up_data_N$Month.N <- as.numeric( format(Up_data_N$Date, format="%m") ) #month
#Sub_males$Month.N <- as.numeric( format(Sub_males$Date, format="%m") ) #month

max.molt = 4
# molt expected around July - September
Pre_molt_juv <- subset(Up_data_N, CW <= 139 & Month.N %in% c(1:max.molt) ) #juveniles moving into sub-adults
Pre_molt_sub <- subset(Up_data_N, CW %in% c(140:160) & Month.N %in% c(1:max.molt))
# error term 
Pre_molt_leg <- subset(Up_data_N, CW >= 161 & Month.N %in% c(1:max.molt))
Data_2020 <- subset(Up_data_N, CW >= 161 & Month.N %in% c(1:max.molt) & Year ==2020)

sim<- NULL #creates empty lists to store the sim for both sub & juvs
n= 1000
molt_inc <- function(data, sex){
  post_molt <- subset(data, Sex == sex, Date : Sex) #ability ot change sex in fxn
  sim <- as.data.frame(post_molt$CW) # sets up data frame w/ initial CW
for(i in 1:n){
   e <- rnorm(1, mean= 0, sd= 3.29) #randomizing agent for vv 
 
   #equation 9 from Zhang et al 2004 & Smith and Jamieson 1989 (BC)
   ifelse(post_molt$Sex == 1,
    sim[,i+1] <- (1.069* sim[1]) + 18.07 + e, 
   ifelse(post_molt$Sex == 2,
     sim[,i+1] <- (0.864* sim[1]) + 32.35 + e, NA))

}
  
   for(j in 1:nrow(sim)){
    x<- unlist(sim[j, c(2:(n+1))])
    post_molt$POST.carapace[j] <- mean(x)
    post_molt$SD[j] <- sqrt(var(x))

  post_molt$diff <- post_molt$POST.carapace - post_molt$CW    
 post_molt$perc <- post_molt$diff / post_molt$POST.carapace
   }
  return(post_molt)
}

post_molt_juv <- molt_inc(Pre_molt_juv, 1)
post_molt_sub <- molt_inc(Pre_molt_sub, 1)
post_molt_leg <- molt_inc(Pre_molt_leg, 1)
post_2020 <- molt_inc(Data_2020, 1) #Use to show direct relationship for 2020 

#H0_1 = diff sub-adults btw April and Sept = 0 (b/c juv cohort moves into  sub-adult cohort)
  #HA_1 != 0
#t.test(Carapace.Width ~ Month, model, var.equal = FALSE)
#var.test(Carapace.Width~ Month, model)
```



# Probability of Molt Calculation (Legal Males)
```{r}
# Before Molt
t_1 = (31 + 28 + 31 +30 + 31 +1) # AKA June 1st in julian dates
# After Molt
t1 = (t_1 + 30 + 31 + 31 + 30 ) # AKA October 1st in julian 
# Instantaneous Molt time 
t0 = t_1 + ( (t1 - t_1)/2) #AKA August 1st

# Equation for CPUE at t0
CPUE.t0 <- rep(NA, nrow(avgs))
for(i in 2:nrow(Est_Mort_legal)){ 
  if(Est_Mort_legal$Year[i] == Est_Mort_legal$Year[i-1] & Est_Mort_legal$Month[i-1] >= 5 & Est_Mort_legal$Month[i] != Est_Mort_legal$Month[i-1] ) #added >= 5 because need it to be close to T-1 (June) so >= May
    {CPUE.t0[i] <- Est_Mort_legal$CPUE[i-1] * exp( -1*Mean.Mort_legal * (t0 - t_1)) }
}
CPUE.t0 <- na.omit(CPUE.t0 ) #2013,2014, 2015, 2020 left
P.molt <- cbind(avgs, CPUE.t0)  #adds the CPUE.t0    
 
#Equation for CPUE.old [old shell ind.]
with (Est_Mort_legal, table(Year, Month)) #when last sample collected by year
CPUE.old <- rep(NA, nrow(Est_Mort_legal))
for(i in 2:nrow(Est_Mort_legal)){
  if(Est_Mort_legal$Year[i] == Est_Mort_legal$Year[i-1] & Est_Mort_legal$Month[i] >= 9 & Est_Mort_legal$Month[i] != Est_Mort_legal$Month[i-1] ) #added >= 9 because need it to be close to T1 (October) so >= September
    {CPUE.old[i] <- Est_Mort_legal$CPUE[i] * exp( -1*Mean.Mort_legal * (t0 - t1)) }
}
CPUE.old <- c(na.omit(CPUE.old ),0) #NOT enough old shell sample days - ONLY 2013, 2014 (until 2020 is updated)

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~REMOVE THE SECOND 0^^ WHEN 2020 GETS UPDATD~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

P.molt <- P.molt[c(1:2,4), ]#removes yeare 2015 because does NOT have sufficient end month (T1)
P.molt <- cbind(P.molt , CPUE.old)  #adds the CPUE.t0    
Prob.Molt_legal <- (P.molt$CPUE.t0 - P.molt$CPUE.old) / P.molt$CPUE.t0
Prob.Molt_legal
 
Mortality <- cbind(Prob.Molt_sub, Prob.Molt_legal)
Mortality <- rbind(Mortality, cbind(Mean.Mort_sub, Mean.Mort_legal))
rownames(Mortality) <- c("2013", "2014", "2020", "Mean.Mort")
Mortality 
```



# DATA FROM MOUTH OF COL RIVER GORGE
## Estimating mortality and prob of molt
```{r}
Sub_males_MCR <- subset(data_MCR, SEX== 1 & CW >= 140 & CW < 159, select = c(Date, Pot.., CW) )

#Sets Months in Order
Sub_males_MCR <- Sub_males_MCR %>% 
  mutate(Month = factor(
    month(Date, label = FALSE),   # thing you're converting
    1:12,                                  # values it could take
    labels =                               # how they should appear
      c("January", "February", "March", "April",
      "May", "June", "July", "August",
      "September", "October", "November", "December"),
    ordered = TRUE)) #Sets the Month to automatically order via month order [lubridate package]

#Adds soak times (Email corespondance 8/20)
Sub_males_MCR$Soak.Time.hr <- ifelse(Sub_males_MCR$Date == "2018-08-23", 23, 
       ifelse(Sub_males_MCR$Date == "2018-08-30", 19,
       ifelse(Sub_males_MCR$Date == "2018-09-18", 24,
       ifelse(Sub_males_MCR$Date == "2018-09-28", 2.5,NA ) ) ) )
#Sort the dataframe by Month for Plot
index <- with(Sub_males_MCR, order(Date, Pot..)) #places the order
Sub_males_MCR <- Sub_males_MCR[index,] #outputs the data in that order

# Making a Julian date by the year 
Sub_males_MCR$J.date <- julian.Date(Sub_males_MCR$Date, as.Date("2018-01-01"), by= 'Year' )


#Prep the dataframe
Sub_males_MCR$Total.pots <- rep(NA,nrow(Sub_males_MCR))
Sub_males_MCR$Total.crabs <- rep(NA,nrow(Sub_males_MCR))
Sub_males_MCR$Total.pots[length(Sub_males_MCR$Total.pots)] <- Sub_males_MCR$Pot[length(Sub_males_MCR$Total.pots)] 

#When there is a different sample date (i.e. sample), records total pots used and number of crabs collected 
j=1
for(i in 2:nrow(Sub_males_MCR)){
  if(Sub_males_MCR$Date[i] !=  Sub_males_MCR$Date[i-1])
  { Sub_males_MCR$Total.pots[j:i] <- Sub_males_MCR$Pot[i-1];
    Sub_males_MCR$Total.crabs[j:i] <- (i-j);
  j=i}
  
  if(i == nrow(Sub_males_MCR)) 
  {Sub_males_MCR$Total.pots[j:i] <- Sub_males_MCR$Pot[i];
  Sub_males_MCR$Total.crabs[j:i] <- i-j +1}
}
Sub_males_MCR <- Sub_males_MCR[ ,c(1,4,6,5,3,2,7,8)]


# EFFORT: = soak time * # of traps in a sample string
Sub_males_MCR$Effort <- Sub_males_MCR$Soak.Time.hr * Sub_males_MCR$Total.pots

# CPUE = # of crabs caught in a string of traps / standard effort
Sub_males_MCR$CPUE <- Sub_males_MCR$Total.crabs / Sub_males_MCR$Effort


#MORTALITY ESTIMATE
Mort_sub_MCR <- aggregate(cbind(J.date, Soak.Time.hr, Total.pots, Effort, CPUE) ~ Date, data = Sub_males_MCR, mean) #reports just the individual date and the CPUE
Mort_sub_MCR$Month <- as.numeric( format(Mort_sub_MCR$Date, format="%m") ) #month
Mort_sub_MCR <- Mort_sub_MCR[ , c(1,7,2:6)]

#estimating mortality for samples in same year but different months
Mort_sub_MCR$N.Mort <- rep(NA, nrow(Mort_sub_MCR)) #NULL column
Mort_sub_MCR$N.Mort[3] <- (-1/( (Mort_sub_MCR$J.date[3]) - (Mort_sub_MCR$J.date[1]) ) )* log(Mort_sub_MCR$CPUE[3] / Mort_sub_MCR$CPUE[1] )  ## USED the 3rd value because low effort due to short soak time in last sample
Mort_sub_MCR


#estimating mort prob. based off of zhang's equation(figure 3)
for(i in 1:nrow(Sub_males_MCR)){
  if(Sub_males_MCR$Carapace.Width[i] <= 150)
  { Sub_males_MCR$Molt.prob[i] <- (-0.0014* Sub_males_MCR$Carapace.Width[i]) + 1.14 }

  if(Sub_males_MCR$Carapace.Width[i] > 150)
  { Sub_males_MCR$Molt.prob[i] <- (-0.014* Sub_males_MCR$Carapace.Width[i]) + 2.71 }
}

ggplot(Sub_males_MCR, aes(x=CW, y= Molt.prob)) +
     geom_point(aes(color = CW)) +
      scale_color_viridis(option = "D")+
    labs(title= "Size-specific molting probabilities - MCR")

```


# Comparing proportions of legal to Sub-legal males 
```{r}
x <- as.data.frame(as.table(with(Up_data_N, table(Sex, Date)) )) #legal males & females
     males <- subset(x, Sex == 1, Date:Freq)  
y <-  as.data.frame(as.table(with(Sub_males, table(Sex, Date)) ))#sub-legal males 

 males <- merge(males, y, by = "Date", all= TRUE)
males <- males[ ,c(1,2,4)] 
 colnames(males) <- c("Date", "Freq.Total", "Freq.sub")
 males$Prop <- males$Freq.sub / males$Freq.Total
 males
 # summary(males$Prop) #average proportion of sub-legal males from total males is 22%
```






# CHECK RAREFACTION FILE AND IF WORKING THEN REMOVE 
# rarefaction of six pots [EXAMPLE of how I was running it before function]
```{r}
sub_sample_6 <- subset(crabs.per.pot, Pot == xyz[[6]][1] | Pot == xyz[[6]][2] | Pot == xyz[[6]][3] |  Pot == xyz[[6]][4] | Pot == xyz[[6]][5] | Pot == xyz[[6]][6])

# Making a Julian date by the year 
sub_sample_6$J.date <- julian.Date(sub_sample_6$Date, as.Date("2020-01-01"), by= 'Year' )

#When there is a different sample date (i.e. sample), records total pots used and number of crabs collected 
sub_sample_6$Total.crabs <- rep(NA,nrow(sub_sample_6))
j=1
for(i in 2:nrow(sub_sample_6)){
  if(sub_sample_6$Date[i] !=  sub_sample_6$Date[i-1])
  { sub_sample_6$Total.crabs[j:i] <- (i-j);
  j=i}

  if(i == nrow(sub_sample_6)) {
    sub_sample_6$Total.crabs[j:i] <- i-j +1
  } }
sub_sample_6$Total.pots <- rep(6, nrow(sub_sample_6))

# EFFORT: = soak time * # of traps in a sample string
sub_sample_6$Effort <- sub_sample_6$Soak.Time.hr * sub_sample_6$Total.pots

# CPUE = # of crabs caught in a string of traps / standard effort
sub_sample_6$CPUE <- sub_sample_6$Total.crabs / sub_sample_6$Effort

#MORTALITY ESTIMATE
Est_sample_6<- aggregate(cbind(J.date, Soak.Time.hr, Total.pots,Total.crabs, Effort, CPUE) ~ Date, data = sub_sample_6, mean) #reports just the individual date and the CPUE
Est_sample_6$Month <- as.numeric( format(Est_sample_6$Date, format="%m") ) #month


#estimating mortality for samples in same year but different months
Est_sample_6$N.Mort <- rep(NA, nrow(Est_sample_6)) #NULL column
for(i in 2: length(Est_sample_6)){
  Est_sample_6$N.Mort[i] <- (-1/( (Est_sample_6$J.date[i]) - (Est_sample_6$J.date[i-1]) ) )*log(Est_sample_6$CPUE[i] / Est_sample_6$CPUE[i-1] )  ## USED the 3rd value because low effort due to short soak time in last sample
}
Est_sample_6
```


# Rarefaction of Sub-adult Males
```{r}
# ONE pot used
Est_1 <- rarefaction(crabs.per.pot.sub.males, 1, xyz[[1]])

for(i in 2: nrow(Est_1)){
  Est_1$N.Mort[i] <- (-1/( (Est_1$J.date[i]) - (Est_1$J.date[i-1]) ) )*log(Est_1$CPUE[i] / Est_1$CPUE[i-1] )  ## USED the 3rd value because low effort due to short soak time in last sample
  }


# TWO pots used
Est_2 <- rarefaction(crabs.per.pot.sub.males, 2, xyz[[2]])

for(i in 2: nrow(Est_2)){
  Est_2$N.Mort[i] <- (-1/( (Est_2$J.date[i]) - (Est_2$J.date[i-1]) ) )*log(Est_2$CPUE[i] / Est_2$CPUE[i-1] )  ## USED the 3rd value because low effort due to short soak time in last sample
  }



# THREE pots used
Est_3 <- rarefaction(crabs.per.pot.sub.males, 3, xyz[[3]])

for(i in 2: nrow(Est_3)){
  Est_3$N.Mort[i] <- (-1/( (Est_3$J.date[i]) - (Est_3$J.date[i-1]) ) )*log(Est_3$CPUE[i] / Est_3$CPUE[i-1] )  ## USED the 3rd value because low effort due to short soak time in last sample
}


# FOUR pots used
Est_4 <- rarefaction(crabs.per.pot.sub.males, 4, xyz[[4]])

for(i in 2: nrow(Est_4)){
  Est_4$N.Mort[i] <- (-1/( (Est_4$J.date[i]) - (Est_4$J.date[i-1]) ) )*log(Est_4$CPUE[i] / Est_4$CPUE[i-1] )  ## USED the 3rd value because low effort due to short soak time in last sample
}


# FIVE pots used
Est_5 <- rarefaction(crabs.per.pot.sub.males, 5, xyz[[5]])

for(i in 2: nrow(Est_5)){
  Est_5$N.Mort[i] <- (-1/( (Est_5$J.date[i]) - (Est_5$J.date[i-1]) ) )*log(Est_5$CPUE[i] / Est_5$CPUE[i-1] )  ## USED the 3rd value because low effort due to short soak time in last sample
}


# SIX pots used
Est_6 <- rarefaction(crabs.per.pot.sub.males, 6, xyz[[6]])

for(i in 2: nrow(Est_6)){
  Est_6$N.Mort[i] <- (-1/( (Est_6$J.date[i]) - (Est_6$J.date[i-1]) ) )*log(Est_6$CPUE[i] / Est_6$CPUE[i-1] )  ## USED the 3rd value because low effort due to short soak time in last sample
}

# SET UP DATAFRAME OF RESULTS
Plot2 <- data.frame(Pots = c(1:6), n.Mort = NA)
  Plot2$n.Mort[1] <- Est_1$N.Mort[nrow(Est_1)]
    #Est_sample_1$N.Mort[nrow(Est_sample_1)] #grabs the last natural mort. rate btw june-july 2020
  Plot2$n.Mort[2] <- Est_2$N.Mort[nrow(Est_2)]
  Plot2$n.Mort[3] <- Est_3$N.Mort[nrow(Est_3)]
  Plot2$n.Mort[4] <- Est_4$N.Mort[nrow(Est_4)]
  Plot2$n.Mort[5] <- Est_5$N.Mort[nrow(Est_5)]
  Plot2$n.Mort[6] <- Est_6$N.Mort[nrow(Est_6)]
  #Plot$Avg.N.Mort <- rep(0.01189,6) # sub adults


 ggplot(Plot2, aes(x=Pots, y=n.Mort))+
    geom_smooth(method = lm, se = T)+
    geom_point()+
    geom_hline(yintercept=0.01189, color= 'red', linetype="dashed")+ #sub adult males
    #geom_hline(yintercept=0.023, color= 'red', linetype="dashed")+ #adults
    guides(color="legend")+
    ggtitle("Rarefaction of Sub-adult Males") 

```